---
title: "Bayesian Data Analysis:  Notes and Exercises"
author: "Lee Przybylski"
date: "1/1/2023"
output:
  html_document:
    toc: true
    toc_float: true
    #extra_dependencies: ["amsthm", "amsmath", "enumerate"]
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Introduction

Below is a write-up of notes and solutions to the exercises from the third 
edition of *Bayesian Data Analysis* by Andrew Gelman et al.  

## Ch 1:  Probability and Inference

### Notes

Here we include some key ideas from the chapter just for reference.

### Exercises

1. Conditional probability:  suppose that if $\theta = 1$, then $y$ has a normal distribution with mean 1 and standard deviation $\sigma$, and if $\theta = 2$, then $y$ has a normal distribution with mean 2 and standard deviation $\sigma$.  Also suppose $\Pr(\theta = 1) = 0.5$ and $\Pr(\theta = 2) = 0.5.$

    (a). For $\sigma = 2$, write the formula for the marginal probability density for $y$ and sketch it.
    
    Solution:  Solution to part (a)
  
    (b). What is $\Pr(\theta = 1|y = 1)$, again, supposing $\sigma = 2$?
    
    Soution:  Solution to part (b)
  
2. Problem 2

## Ch 2: Single Parameter Models

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
